On the uniqueness and stability of dictionaries for sparse representation of noisy signals

### 2017-04-10 manuscripts/PNAS/chazthm_pnas.tex
SR Lem 4

\begin{lemma}\label{NonEmptyLemma} 
Fix $\ell, m$, and $\bar m$. Suppose $E \subseteq 2^{[m]}$ is $\ell$-regular with the SIP.  If $\pi: E \to 2^{[\bar m]}$ has $\sum_{S \in E} |S| = \sum_{S \in E} |\pi(S)|$ and:
\begin{align}\label{cond}
|\bigcap_{S \in F} \pi(S)| \leq |\bigcap_{S \in F} S |,\ \ \   \text{for } F \subseteq E,
\end{align}
then $\bar m \geq m$ and the association $i \mapsto \cap \pi(F(i))$ defines an injective map from $J$ to $[\bar m]$ for some $J \subseteq [m]$ of size $p = \bar m - \ell(\bar m - m)$. In particular, if $\bar m = m$ then the map $\pi$ is induced by a permutation.
\end{lemma}

\begin{proof}
Consider the collection of pairs: \[T_1 = \{(i, S): i \in \pi(S), S \in E\},\] which number $|T_1| = \sum_{S \in E} |\pi(S)| = \sum_{S \in E} |S| = \ell m$ by regularity of $E$.  
We first show that $\bar m \geq m$. Otherwise, if $\bar m < m$, then by pigeonholing the $m \ell \geq \ell (\bar m + 1) \geq [(\ell + 1) - 1] \bar m + 1$ elements of $T_1$ with respect to the set $[\bar m]$ of possible first indices, we see that there must exist at least $\ell + 1$ elements of $T_1$ having the same first index. But by \eqref{cond} there can be no more than $\ell$ elements of $T_1$ with a given first index, since $E$ is $\ell$-regular. Hence $\bar m \geq m$.

Next, since \[|T_1| > \bar m (\ell -1) + (p-1) > \bar m (\ell -1),\] the pigeonhole principle also implies that there are at least $\ell$ elements of $T_1$ sharing the same first coordinate. Let $F_1$ be the preimage of this family under $\pi$.  By the assumption \eqref{cond}, the intersection of $F_1$ is nonempty, and in fact must contain exactly $\ell$ elements. Thus, $F_1$ is the star of some element $i_1 \in [m]$.  From the SIP, the intersection of sets in $F_1$ is $\{i_1\}$, which by \eqref{cond} implies that the intersection of $\pi(S)$ with $S \in F_1$ also contains exactly one element.  Set $T_2 = T_1 \setminus \{(i,S) \in T_1: i = i_1\}$, which contains at least $(\bar m - 1)(\ell -1)$ sets. Iterating this argument $p = \bar{m} - \ell(\bar m - m)$ times produces the desired $J = \{i_1, \ldots, i_p\} \subseteq [m]$.
\end{proof}
\begin{proof}[Proof of Lem.\ref{MainLemma}]
We begin by showing that: 
\begin{align}\label{eq2}
d(\text{\rm Span}\{\mathbf{B}_{\pi(S)}\}, \text{\rm Span}\{\mathbf{A}_S\} ) = d(\text{\rm Span}\{\mathbf{A}_S\}, \text{\rm Span}\{\mathbf{B}_{\pi(S)}\}).
\end{align}
Since the right-hand side of \eqref{GapUpperBound} is strictly less than one (by \eqref{delrho} and $r \leq 1$), it follows by \eqref{dimLem} that $|\pi(S)| \geq \dim(\text{\rm Span}\{\mathbf{B}_{\pi(S)}\}) \geq \dim(\text{\rm Span}\{\mathbf{A}_{S}\}) = |S|$, with the equality due to $\mathbf{A}$ satisfying \eqref{SparkCondition}. Since $|S| = |\pi(S)|$, we therefore have $\dim(\text{\rm Span}\{\mathbf{B}_{\pi(S)}\}) = \dim(\text{\rm Span}\{\mathbf{A}_{S}\})$, and \eqref{eq2} follows by \eqref{eqdim}. Note that the columns of $\mathbf{B}_{\pi(S)}$ are therefore linearly independent, for all $S \in E$.

We next verify \eqref{cond} holds.  Fix $F \subseteq E$. Since $\text{\rm Span}\{\mathbf{B}_{\cap_{S \in F}\pi(S)}\} \subseteq \cap_{S \in F} \text{\rm Span}\{\mathbf{B}_{\pi(S)}\}$, if $\cap_{S \in F} \text{\rm Span}\{\mathbf{B}_{\pi(S)}\} = \textbf{0}$ then we must have $|\cap_{S \in F} \pi(S)| = 0$ (as the columns of $\mathbf{B}_{\pi(S)}$ are linearly independent) and \eqref{cond} trivially holds. Suppose then that the intersection is not the zero vector. By Lem.~\ref{SpanIntersectionLemma} and Lem.~\ref{DistanceToIntersectionLemma}, and incorporating \eqref{eq2} and \eqref{GapUpperBound}, we have:
\begin{align}\label{randoml}
d( \text{\rm Span}&\{\mathbf{B}_{\cap_{S \in F}\pi(S)}\}, \text{\rm Span}\{\mathbf{A}_{\cap_{S \in F} S}\}  ) \nonumber \\
&\leq d\left( \cap_{S \in F} \text{\rm Span}\{\mathbf{B}_{\pi(S)}\}, \cap_{S \in F} \text{\rm Span}\{\mathbf{A}_{S}\} \right) \nonumber \\
&\leq \sum_{T \in F} \frac{ d\left( \cap_{S \in F} \text{\rm Span}\{\mathbf{B}_{\pi(S)}\},\text{\rm Span}\{\mathbf{A}_{T}\} \right) }{ r( \{ \text{\rm Span}\{\mathbf{A}_{S}\}_{S \in F}) } \nonumber \\
&\leq \sum_{T \in F} \frac{ d\left( \text{\rm Span}\{\mathbf{B}_{\pi(T)}\},\text{\rm Span}\{\mathbf{A}_{T}\} \right) }{ r( \{ \text{\rm Span}\{\mathbf{A}_{S}\}_{S \in F}) }\nonumber \\
&\leq \frac{|F| \varepsilon}{r( \{ \text{\rm Span}\{\mathbf{A}_{S}\} \}_{S \in F})} 
\leq \frac{C_2 \varepsilon}{\max_i\|\mathbf{A}_i\|_2}. 
\end{align}
Since $\varepsilon < L_2(\mathbf{A}) / C_2$, by \eqref{delrho} the right-hand side in \eqref{randoml} is strictly less than one. Hence, $\dim(\text{\rm Span}\{\mathbf{B}_{\cap_{S \in F}\pi(S)}\}) \leq \dim(\text{\rm Span}\{\mathbf{A}_{\cap_{S \in F} S}\})$ by \eqref{dimLem}, and \eqref{cond} follows from the linear independence of the columns of $\mathbf{A}_{S}$ and $\mathbf{B}_{\pi(S)}$ for all $S \in F$.

>>>>

%\begin{lemma}\label{NonEmptyLemma} 
%Fix $\ell, m$, and $\bar m$. Suppose $E \subseteq 2^{[m]}$ is $\ell$-regular with the SIP.  If $\pi: E \to 2^{[\bar m]}$ has $\sum_{S \in E} |S| = \sum_{S \in E} |\pi(S)|$ and:
%\begin{align}\label{cond}
%|\bigcap_{S \in F} \pi(S)| \leq |\bigcap_{S \in F} S |,\ \ \   \text{for } F \subseteq E,
%\end{align}
%then $\bar m \geq m$ and the association $i \mapsto \cap \pi(F(i))$ defines an injective map from $J$ to $[\bar m]$ for some $J \subseteq [m]$ of size $p = \bar m - \ell(\bar m - m)$. In particular, if $\bar m = m$ then the map $\pi$ is induced by a permutation.
%\end{lemma}
% i'm placing this on hold since one thing i found was that i needed pi to be injective, but for that i needed the fact that |pi(S)| = |S|

\begin{lemma}\label{NonEmptyLemma} 
Fix $\ell, m$, and $\bar m$. Suppose $E \subseteq 2^{[m]}$ is $\ell$-regular with the SIP.  If $\pi: E \to 2^{[\bar m]}$ has $|S| = |\pi(S)|$ and:
\begin{align}\label{cond}
|\bigcap_{S \in F} \pi(S)| \leq |\bigcap_{S \in F} S |,\ \ \   \text{for } F \subseteq E,
\end{align}
then $\bar m \geq m$ and the association $i \mapsto \cap \pi(F(i))$ defines an injective map from $J$ to $[\bar m]$ for some $J \subseteq [m]$ of size $p = \bar m - \ell(\bar m - m)$. In particular, if $\bar m = m$ then the map $\pi$ is induced by a permutation.
\end{lemma}

\begin{proof}
First observe that assumption \eqref{cond} already constrains $\pi$ to be injective:  if $S_1 \neq S_2$, then $|\pi(S_1) \cap \pi(S_2)| \leq |S_1 \cap S_2| < \max\{|S_1|, |S_2|\} = \max\{|\pi(S_1)|, |\pi(S_2)|\}$, which implies that $\pi(S_1) \neq \pi(S_2)$.  Next, consider the collection of pairs: \[T_1 := \{(i, S): i \in \pi(S), S \in E\},\] which number $|T_1| = \sum_{S \in E} |\pi(S)| = \sum_{S \in E} |S| = \ell m$ by regularity of $E$.  
Our second observation is that $\bar m \geq m$. Otherwise, if $\bar m < m$, then by pigeonholing the $m \ell \geq \ell (\bar m + 1) \geq [(\ell + 1) - 1] \bar m + 1$ elements of $T_1$ with respect to the set $[\bar m]$ of possible first indices, there must exist at least $\ell + 1$ elements of $T_1$ having the same first index. But by \eqref{cond} and injectivity of $\pi$, we then have more than $\ell$ edges in $E$ with common element, a contradiction to regularity.

Since $|T_1| > \bar m (\ell -1) + (p-1) \geq \bar m (\ell -1)$, the pigeonhole principle also implies that there are at least $\ell$ elements of $T_1$ sharing the same first coordinate. Let $F_1$ be the preimage of this family under $\pi$.  By the assumption \eqref{cond}, the intersection of $F_1$ is nonempty, and in fact must contain exactly $\ell$ elements. Thus, $F_1$ is the star of some element $i_1 \in [m]$.  From the SIP, the intersection of sets in $F_1$ is $\{i_1\}$, which by \eqref{cond} implies that the intersection of $\pi(S)$ with $S \in F_1$ also contains exactly one element.  If $p > 1$, then define $T_2 := T_1 \setminus \{(i,S) \in T_1: i = i_1\}$, which contains more than $(\bar m - 1)(\ell -1)$ sets, and produce by the same argument a (necessarily) different index $i_2$. Iterating $p = \bar{m} - \ell(\bar m - m)$ times this loop yields the desired $J = \{i_1, \ldots, i_p\} \subseteq [m]$.
\end{proof}
\begin{proof}[Proof of Lem.\ref{MainLemma}]
We begin by showing that: 
\begin{align}\label{eq2}
d(\text{\rm Span}\{\mathbf{B}_{\pi(S)}\}, \text{\rm Span}\{\mathbf{A}_S\} ) = d(\text{\rm Span}\{\mathbf{A}_S\}, \text{\rm Span}\{\mathbf{B}_{\pi(S)}\}).
\end{align}
Since the right-hand side of \eqref{GapUpperBound} is strictly less than one (by \eqref{delrho} and $r \leq 1$), it follows by \eqref{dimLem} that $|\pi(S)| \geq \dim(\text{\rm Span}\{\mathbf{B}_{\pi(S)}\}) \geq \dim(\text{\rm Span}\{\mathbf{A}_{S}\}) = |S|$, with the equality due to $\mathbf{A}$ satisfying \eqref{SparkCondition}. Since $|S| = |\pi(S)|$, we therefore have $\dim(\text{\rm Span}\{\mathbf{B}_{\pi(S)}\}) = \dim(\text{\rm Span}\{\mathbf{A}_{S}\})$, and \eqref{eq2} follows by \eqref{eqdim}. Note that the columns of $\mathbf{B}_{\pi(S)}$ are therefore linearly independent, for all $S \in E$.

We next verify \eqref{cond} holds.  Fix $F \subseteq E$. Since $\text{\rm Span}\{\mathbf{B}_{\cap_{S \in F}\pi(S)}\} \subseteq \cap_{S \in F} \text{\rm Span}\{\mathbf{B}_{\pi(S)}\}$, if $\cap_{S \in F} \text{\rm Span}\{\mathbf{B}_{\pi(S)}\} = \{\textbf{0}\}$ then we must have $|\cap_{S \in F} \pi(S)| = 0$ (as the columns of $\mathbf{B}_{\pi(S)}$ are linearly independent) and \eqref{cond} trivially holds. Suppose then that the intersection is not the zero vector. By Lem.~\ref{SpanIntersectionLemma} and Lem.~\ref{DistanceToIntersectionLemma}, and incorporating \eqref{eq2} and \eqref{GapUpperBound}, we have:
\begin{align}\label{randoml}
d( \text{\rm Span}&\{\mathbf{B}_{\cap_{S \in F}\pi(S)}\}, \text{\rm Span}\{\mathbf{A}_{\cap_{S \in F} S}\}  ) \nonumber \\
&\leq d\left( \cap_{S \in F} \text{\rm Span}\{\mathbf{B}_{\pi(S)}\}, \cap_{S \in F} \text{\rm Span}\{\mathbf{A}_{S}\} \right) \nonumber \\
&\leq \sum_{T \in F} \frac{ d\left( \cap_{S \in F} \text{\rm Span}\{\mathbf{B}_{\pi(S)}\},\text{\rm Span}\{\mathbf{A}_{T}\} \right) }{ r( \{ \text{\rm Span}\{\mathbf{A}_{S}\}_{S \in F}) } \nonumber \\
&\leq \sum_{T \in F} \frac{ d\left( \text{\rm Span}\{\mathbf{B}_{\pi(T)}\},\text{\rm Span}\{\mathbf{A}_{T}\} \right) }{ r( \{ \text{\rm Span}\{\mathbf{A}_{S}\}_{S \in F}) }\nonumber \\
&\leq \frac{|F| \varepsilon}{r( \{ \text{\rm Span}\{\mathbf{A}_{S}\} \}_{S \in F})} 
\leq \frac{C_2 \varepsilon}{\max_i\|\mathbf{A}_i\|_2}. 
\end{align}
Since $\varepsilon < L_2(\mathbf{A}) / C_2$, by \eqref{delrho} the right-hand side in \eqref{randoml} is strictly less than one. Hence, $\dim(\text{\rm Span}\{\mathbf{B}_{\cap_{S \in F}\pi(S)}\}) \leq \dim(\text{\rm Span}\{\mathbf{A}_{\cap_{S \in F} S}\})$ by \eqref{dimLem}, and \eqref{cond} follows from the linear independence of the columns of $\mathbf{A}_{S}$ and $\mathbf{B}_{\pi(S)}$ for all $S \in F$.
